Summary of the things that I learned

### 2024-12-09 700-formalization.tex
Improve notation and wording

Since Hyland's paper uses a lot of general category theory, formalizing it entailed adding the preliminaries that had not yet been formalized to the library. Among these are:
\begin{itemize}
  \item Univalence of the Sigma displayed category (\coqident{CategoryTheory.DisplayedCats.Examples.Sigma}{is_univalent_sigma_disp}).
  \item The Sigma displayed category creates limits (\coqident{CategoryTheory.DisplayedCats.Examples.Sigma}{creates_limits_sigma_disp_cat}).
  \item The definitions $ \overline \C $ (\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_envelope}) and $ \tilde \C $ (\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_envelope'}) for the Karoubi envelope, together with some properties and their equivalence (\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_equivalence}).
  \item The contents of Section \ref{sec:monoid-category}: The functorial construction of a one-object category from a monoid (\coqident{CategoryTheory.Categories.MonoidToCategory}{monoid_to_category}), the equivalence between it's presheaves and sets with a right monoid action (\coqident{CategoryTheory.Categories.MonoidToCategory}{monoid_presheaf_action_equivalence}), properties of the category of sets with a monoid action and restriction and extension of scalars (\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.MonoidActions}{scalar_restriction_functor}, \coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.MonoidActions}{scalar_extension_functor}).
  \item The Yoneda embedding preserves exponential objects (\coqident{CategoryTheory.YonedaExponentials}{yoneda_preserves_exponentials}).
  \item The uniqueness of the Rezk completion (\coqident{Bicategories.RezkCompletions.Uniqueness}{rezk_completion_unique}).
% FundamentalTheorem/SurjectivePrecomposition.v
  \item If $ F : \C_1 \to \C_2 $ is a fully faithful functor, and $ \D $ is a category with colimits, then the precomposition functor $ F \bullet - : [\C_2, \D] \to [\C_1, \D] $ is split essentially surjective (\coqident{AlgebraicTheories.FundamentalTheorem.SurjectivePrecomposition}{pre_comp_split_essentially_surjective}).
  \item A proof of a generalized version of Lemma \ref{lem:equivalent-presheaf-cats}, showing that if $ F_1 : \C_1 \to \C_2 $ is a functor, if $ F_2 : \C_2 \to \C_3 $ is a fully faithful functor, if $ \D $ is a category with colimits and if the precomposition $ (F_1 \bullet F_2) \bullet - : [\C_3, \D] \to [\C_1, \D] $ is an adjoint equivalence, then $ F_1 \bullet - $ and $ F_2 \bullet - $ are adjoint equivalences too (\coqident{CategoryTheory.Equivalences.EquivalenceFromComp}{adjoint_equivalence_1_from_comp}).
% IndexedSetCategory.v
  \item The univalent category $ \SET^A $ of indexed sets $ (X_a)_a $ over a type $ A $ (\coqident{AlgebraicTheories.IndexedSetCategory}{indexed_set_cat}), which can be defined formally as
  \[ \SET^A_0 = (A \to \SET) \quad \text{and} \quad \SET^A(X, Y) = \prod_a X_a \to Y_a \]
  that is univalent (\coqident{AlgebraicTheories.IndexedSetCategory}{is_univalent_indexed_set_cat}) and has limits (\coqident{AlgebraicTheories.IndexedSetCategory}{limits_indexed_set_cat}).
\end{itemize}

@@

A displayed category $ D $ over a category $ C $ firstly consists of a type $ D_c $ for every $ c : C $ (corresponding to the type of objects $ F^{-1}(c) $) and a type $ d \to_f d^\prime $ for all $ d: D_c $, $ d^\prime : D_{c^\prime} $ and $ f : C(c, c^\prime) $ (corresponding to $ C^\prime((c, d), (c^\prime, d^\prime)) $). A displayed category also consists of an identity `morphism' $ d \to_{\id c} d $ for all $ d : D_c $, and compositions $ \bar f \cdot \bar g : d \to_{f \cdot g} d^{\prime \prime} $ for $ \bar f : d \to_f d^\prime $ and $ \bar g : d^\prime \to_g d^{\prime \prime} $.

@@

One of the places where displayed categories are conceptually better to work with than forgetful functors, is in the case of fibrations. Recall that a functor $ P: C^\prime \to C $ is a fibration if for every $ Y : C^\prime $ and $ f : C(X, P(Y)) $, there exists $ \bar X : C^\prime $ with $ P(\bar X) = X $ (and a cartesian morphism $ \bar f : C^\prime(\bar X, Y) $ with $ P(\bar f) = f $). The equality $ P(\bar X) = X $ is on objects in the category, and this violates the \textit{principle of equivalence}: `if something is true for $ A $, and $ A $ is isomorphic to $ B $, then it should also hold for $ B $'. Of course, if the category is univalent, isomorphism and equality are the same, but definitions that use equality on objects still give a bit of conceptual friction.

However, we can also define this in the language of displayed categories. The definition becomes
\begin{definition}
  A displayed category $ D $ over $ C $ is a fibration if for every $ \bar Y : D Y $ and $ f: C(X, Y) $, we have $ \bar X : D X $ and a cartesian morphism $ \bar f : \bar X \to_f \bar Y $.
\end{definition}
This avoids using equality on objects, because we can just talk about `the objects above $ X $'. Therefore, in UniMath fibrations are defined in this way and the algebraic theories package uses this definition to show that the displayed categories of algebraic theory algebras and presheaves, over the category of algebraic theories, are fibrations.

@@

This is all very reminiscent of the way that displayed categories `borrow' information from their base category. Now if a category $ C $ has binary products, we say that a displayed category $ D $ over $ C $ \textit{creates binary products}, if for all $ \bar X : D_X $, $ \bar Y : D_Y $, we can `lift' their product. That is, if we can find
\[ \overline{X \times Y} : D_{X \times Y}, \quad \bar{\pi_1} : \overline{X \times Y} \to_{\pi_1} \bar X \quad \text{and} \quad \bar{\pi_2} : \overline{X \times Y} \to_{\pi_2} \bar Y \]
and if $ (X \times Y, \overline{X \times Y}) $ with the projections $ (\pi_1, \bar{\pi_1}) $ and $ (\pi_2, \bar{\pi_2}) $ is the product of $ (X, \bar X) $ and $ (Y, \bar Y) $ in $ \int_C D $. By definition, if $ D $ creates binary products, then $ \int_C D $ has binary products.

@@

During this project, I added a small lemma to the library, showing that for a full subcategory $ \int_C D \subseteq C $, where $ D_c $ is a mere proposition for all $ c : C $, if for all $ (X, \bar X), (Y, \bar Y) : \int D $, we have $ D_{X \times Y} $ (the $ C $-product of two objects in $ \int D $ is again in $ \int D $), then $ D $ creates binary products. I also added a lemma showing that if a displayed category $ D $ over $ C $, and a displayed category $ E $ over $ D $, both create binary products, then $ \sigma_D E $ also creates binary products.

Note that for many categories, the displayed morphisms are mere propositions. In particular, this holds for the `\dots data' categories of this project. For example, an algebra morphism is a function that `respects the operation', and respecting the operation is a mere proposition. In these categories, it is not necessary to give a full proof that the object $ (X \times Y, \overline{X \times Y}) $ and the morphisms $ (\pi_i, \overline{\pi_i}) $ form a binary product, because it suffices to show that for all $ C, \bar C, f, g $ and all $ \bar f : \bar C \to_f \bar A $ and $ \bar g : \bar C \to_g \bar B $, we can lift the product morphism to
\[ \overline{\langle f, g \rangle} : \bar C \to_{\langle f, g \rangle} \overline{A \times B} \]
For example, if the function $ f $ commutes with the algebra actions on $ \bar C $ and $ \bar A $, and $ g $ commutes with the actions on $ \bar C $ and $ \bar B $, then we need to show that the product morphism $ \langle f, g \rangle $ commutes with the actions on $ \bar C $ and $ \overline{A \times B} $. I added a lemma showing that it suffices to construct a lift for the product morphism, and used this lemma to show that the `\dots data' displayed categories create binary products.

Now, recall that the categories of $ T $-algebras and $ T $-presheaves are a fiber of a displayed category. Now, if we want a binary product of $ X, Y : \Alg T $ and we take the product of $ (T, X) $ and $ (T, Y) $ in the category of all algebras, we end up with $ (T \times T, X \times Y) $, or $ X \times Y : \Alg {T \times T} $, even though we would like to have $ (T, X) $ for some $ X : \Alg T $. It turns out that here we need the fact that algebras and presheaves are fibered over algebraic theories, and adding the following lemma to the library gave the final brick for showing show that all categories discussed here have limits:
\begin{lemma}
  For a displayed category $ D $ over a category $ C $, suppose that $ C $ has binary products, that $ D $ creates binary products, and that $ D $ is a fibration. Then the fiber categories $ D_X $ have binary products.
\end{lemma}
\begin{proof}
  Take $ \bar X_1, \bar X_2 : D_X $. We have a product $ \bar X_1 \times \bar X_2 : D_{X \times X} $ with projections $ \bar \pi_i : \bar X_1 \times \bar X_2 \to_{\pi_i} \bar X_i $. For the diagonal morphism on $ X $, the fibration gives an object $ \bar Y $ and a a cartesian lift:
  \[ \bar f : \bar Y \to_{\langle \id X, \id X \rangle} \bar X_1 \times \bar X_2, \]
  so we have projections
  \[ \bar p_i = \bar f \cdot \bar \pi_i : \bar Y \to_{\id X} \bar X_i. \]
  In the diagram below, the first row is the fiber $ D_X $ and the second row is the fiber $ D_{X \times X} $.
  \begin{center}
    \begin{tikzcd}
      \bar X_i & \bar Y \arrow[d, "\bar f" description] \arrow[l, "\bar p_i"] & \bar Z \arrow[ll, "\bar g_i"', bend right] \arrow[l, "\bar h"] \arrow[ld, "{\langle \bar g_1, \bar g_2 \rangle}"] \\
      & \bar X_1 \times \bar X_2 \arrow[lu, "\bar \pi_i"] &
    \end{tikzcd}
  \end{center}
  Now, given some $ \bar g_1: \bar Z \to_{g_1} \bar X_1 $ and $ \bar g_2: \bar Z \to_{g_2} \bar X_2 $, we have
  \[ \langle \bar g_1, \bar g_2 \rangle : \bar Z \to_{\langle g_1, g_2 \rangle} \bar X_1 \times \bar X_2. \]
  Because $ \bar f $ is cartesian, we have a unique
  \[ \bar h: \bar Z \to_{\id X} \bar Y, \]
  such that
  \[ \bar h \cdot \bar f = \langle \bar g_1, \bar g_2 \rangle. \]
  Therefore,
  \[ \bar h \cdot \bar p_i = \bar h \cdot \bar f \cdot \bar \pi_i = \langle \bar g_1, \bar g_2 \rangle \cdot \bar \pi_i = \bar g_i, \]
  which shows that $ \bar Y $ is the product of $ \bar X_1 $ and $ \bar X_2 $ in $ D_X $.
\end{proof}

@@

\section{Quotients}\label{sec:quotients}
This project used quotients twice, once explicitly and once implicitly. The first occurrence of quotients was the construction of the `extension of scalars' functor $ f_*: \RAct N \to \RAct M $ from Lemma \ref{lem:scalar-extension}. Here, the formalization worked directly with quotients over a relation. However, the relation given by $ R : (x n, m) \sim (x, f(n) \cdot m) $ for all $ a $, is reflexive and transitive, but not necessarily symmetric. Therefore, we first take the `equivalence closure' $ \bar R $ of $ R $: the smallest transitive and symmetric relation that contains $ R $. This is then used for the quotient. To construct functions from this, like the right action of the monoid $ M^\prime $, we use the universal property of the quotient: a function from $ X \times M^\prime $ that sends `related' elements $ (x_1, m_1) $ and $ (x_2, m_2) $ to the same element gives a function from the quotient $ X \times M^\prime / \sim $. However, the quotient is taken over $ \bar R $ instead of $ R $, so it is no longer possible to assume that $ x_2 = x_1 n $ and $ m_1 = f(n) \cdot m_2 $ for some $ n $. In the case of the right $ M^\prime $-action, it sufficed to add the following lemma to the library
\begin{lemma}
  For a relation $ S $ on a set $ A $ and a relation $ T $ on a set $ B $, if for $ f: A \to B $, we have that $ S(a_1, a_2) $ implies $ T(f(a_1), f(a_2)) $ for all $ a_1, a_2 : A $, then $ \bar S(a_1, a_2) $ implies $ \bar T(f(a_1), f(a_2)) $ for all $ a_1, a_2 $.
\end{lemma}
However, overall, working with these quotients in an elementary way was quite a hassle.

The second occurrence of quotients was in the proof of Corollary \ref{cor:karoubi-presheaf}, or the lemma that it is based on. Initially, the statement of the lemma was only about $ \SET $, stating that $ \iota_C \bullet - : [\bar C, \SET] \to [C, \SET] $ was an adjoint equivalence. The proof used coequalizers of functors, like in Lemma \ref{lem:karoubi-equivalence}. Note that coequalizers of set-valued functions are given pointwise by coequalizers in $ \SET $, which are formalized using quotients. Because of this, when simplifying, rocq usually reduced the goal to elementary statements about quotients, which somewhat obfuscated the relatively simple nature of the construction. Luckily, the construction could be made agnostic to the specific implementation of coequalizers, by introducing a variable
\[ \mathtt{H : Coequalizers HSET} \]
and using this instead of our known implementation. Since the proofs did not use any elements of the sets in question, it was then trivial to generalize to any category with coequalizers. It turned out that the construction was a special version of left Kan extension, which uses general colimits, and using the existing definition of Kan extension, I was able to considerably shorten the proof.

>>>>

Since Hyland's paper uses a lot of general category theory, formalizing it entailed adding the preliminaries that had not yet been formalized to the library. Among these are:
\begin{itemize}
  \item Univalence of the Sigma displayed category (\coqident{CategoryTheory.DisplayedCats.Examples.Sigma}{is_univalent_sigma_disp}).
  \item The Sigma displayed category creates limits (\coqident{CategoryTheory.DisplayedCats.Examples.Sigma}{creates_limits_sigma_disp_cat}).
  \item The definitions $ \overline \C $ (\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_envelope}) and $ \widetilde \C $ (\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_envelope'}) for the Karoubi envelope, together with some properties and their equivalence (\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.KaroubiEnvelope}{karoubi_equivalence}).
  \item The contents of Section \ref{sec:monoid-category}: The functorial construction of a one-object category from a monoid (\coqident{CategoryTheory.Categories.MonoidToCategory}{monoid_to_category}), the equivalence between it's presheaves and sets with a right monoid action (\coqident{CategoryTheory.Categories.MonoidToCategory}{monoid_presheaf_action_equivalence}), properties of the category of sets with a monoid action and restriction and extension of scalars (\coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.MonoidActions}{scalar_restriction_functor}, \coqident{AlgebraicTheories.FundamentalTheorem.CommonUtilities.MonoidActions}{scalar_extension_functor}).
  \item The Yoneda embedding preserves exponential objects (\coqident{CategoryTheory.YonedaExponentials}{yoneda_preserves_exponentials}).
  \item The uniqueness of the Rezk completion (\coqident{Bicategories.RezkCompletions.Uniqueness}{rezk_completion_unique}).
% FundamentalTheorem/SurjectivePrecomposition.v
  \item If $ F : \C_1 \to \C_2 $ is a fully faithful functor, and $ \D $ is a category with colimits, then the precomposition functor $ F \bullet - : [\C_2, \D] \to [\C_1, \D] $ is split essentially surjective (\coqident{AlgebraicTheories.FundamentalTheorem.SurjectivePrecomposition}{pre_comp_split_essentially_surjective}).
  \item A proof of a generalized version of Lemma \ref{lem:equivalent-presheaf-cats}, showing that if $ F_1 : \C_1 \to \C_2 $ is a functor, if $ F_2 : \C_2 \to \C_3 $ is a fully faithful functor, if $ \D $ is a category with colimits and if the precomposition $ (F_1 \bullet F_2) \bullet - : [\C_3, \D] \to [\C_1, \D] $ is an adjoint equivalence, then $ F_1 \bullet - $ and $ F_2 \bullet - $ are adjoint equivalences too (\coqident{CategoryTheory.Equivalences.EquivalenceFromComp}{adjoint_equivalence_1_from_comp}).
% IndexedSetCategory.v
  \item The univalent category $ \SET^A $ of indexed sets $ (X_a)_a $ over a type $ A $ (\coqident{AlgebraicTheories.IndexedSetCategory}{indexed_set_cat}), which can be defined formally as
  \[ \SET^A_0 = (A \to \SET) \quad \text{and} \quad \SET^A(X, Y) = \prod_a X_a \to Y_a \]
  that is univalent (\coqident{AlgebraicTheories.IndexedSetCategory}{is_univalent_indexed_set_cat}) and has limits (\coqident{AlgebraicTheories.IndexedSetCategory}{limits_indexed_set_cat}).
\end{itemize}

@@

A displayed category $ D $ over a category $ C $ firstly consists of a type $ D_c $ for every $ c : C $ (corresponding to the type of objects $ F^{-1}(c) $) and a type $ d \to_f d^\prime $ for all $ d: D_c $, $ d^\prime : D_{c^\prime} $ and $ f : C(c, c^\prime) $ (corresponding to $ C^\prime((c, d), (c^\prime, d^\prime)) $). A displayed category also consists of an identity `morphism' $ d \to_{\id c} d $ for all $ d : D_c $, and compositions $ \overline f \cdot \overline g : d \to_{f \cdot g} d^{\prime \prime} $ for $ \overline f : d \to_f d^\prime $ and $ \overline g : d^\prime \to_g d^{\prime \prime} $.

@@

One of the places where displayed categories are conceptually better to work with than forgetful functors, is in the case of fibrations. Recall that a functor $ P: C^\prime \to C $ is a fibration if for every $ Y : C^\prime $ and $ f : C(X, P(Y)) $, there exists $ \overline X : C^\prime $ with $ P(\overline X) = X $ (and a cartesian morphism $ \overline f : C^\prime(\overline X, Y) $ with $ P(\overline f) = f $). The equality $ P(\overline X) = X $ is on objects in the category, and this violates the \textit{principle of equivalence}: `if something is true for $ A $, and $ A $ is isomorphic to $ B $, then it should also hold for $ B $'. Of course, if the category is univalent, isomorphism and equality are the same, but definitions that use equality on objects still give a bit of conceptual friction.

However, we can also define this in the language of displayed categories. The definition becomes
\begin{definition}
  A displayed category $ D $ over $ C $ is a fibration if for every $ \overline Y : D Y $ and $ f: C(X, Y) $, we have $ \overline X : D X $ and a cartesian morphism $ \overline f : \overline X \to_f \overline Y $.
\end{definition}
This avoids using equality on objects, because we can just talk about `the objects above $ X $'. Therefore, in UniMath fibrations are defined in this way and the algebraic theories package uses this definition to show that the displayed categories of algebraic theory algebras and presheaves, over the category of algebraic theories, are fibrations.

@@

This is all very reminiscent of the way that displayed categories `borrow' information from their base category. Now if a category $ C $ has binary products, we say that a displayed category $ D $ over $ C $ \textit{creates binary products}, if for all $ \overline X : D_X $, $ \overline Y : D_Y $, we can `lift' their product. That is, if we can find
\[ \overline{X \times Y} : D_{X \times Y}, \quad \overline{\pi_1} : \overline{X \times Y} \to_{\pi_1} \overline X \quad \text{and} \quad \overline{\pi_2} : \overline{X \times Y} \to_{\pi_2} \overline Y \]
and if $ (X \times Y, \overline{X \times Y}) $ with the projections $ (\pi_1, \overline{\pi_1}) $ and $ (\pi_2, \overline{\pi_2}) $ is the product of $ (X, \overline X) $ and $ (Y, \overline Y) $ in $ \int_C D $. By definition, if $ D $ creates binary products, then $ \int_C D $ has binary products.

@@

During this project, I added a small lemma to the library, showing that for a full subcategory $ \int_C D \subseteq C $, where $ D_c $ is a mere proposition for all $ c : C $, if for all $ (X, \overline X), (Y, \overline Y) : \int D $, we have $ D_{X \times Y} $ (the $ C $-product of two objects in $ \int D $ is again in $ \int D $), then $ D $ creates binary products. I also added a lemma showing that if a displayed category $ D $ over $ C $, and a displayed category $ E $ over $ D $, both create binary products, then $ \sigma_D E $ also creates binary products.

Note that for many categories, the displayed morphisms are mere propositions. In particular, this holds for the `\dots data' categories of this project. For example, an algebra morphism is a function that `respects the operation', and respecting the operation is a mere proposition. In these categories, it is not necessary to give a full proof that the object $ (X \times Y, \overline{X \times Y}) $ and the morphisms $ (\pi_i, \overline{\pi_i}) $ form a binary product, because it suffices to show that for all $ C, \overline C, f, g $ and all $ \overline f : \overline C \to_f \overline A $ and $ \overline g : \overline C \to_g \overline B $, we can lift the product morphism to
\[ \overline{\langle f, g \rangle} : \overline C \to_{\langle f, g \rangle} \overline{A \times B} \]
For example, if the function $ f $ commutes with the algebra actions on $ \overline C $ and $ \overline A $, and $ g $ commutes with the actions on $ \overline C $ and $ \overline B $, then we need to show that the product morphism $ \langle f, g \rangle $ commutes with the actions on $ \overline C $ and $ \overline{A \times B} $. I added a lemma showing that it suffices to construct a lift for the product morphism, and used this lemma to show that the `\dots data' displayed categories create binary products.

Now, recall that the categories of $ T $-algebras and $ T $-presheaves are a fiber of a displayed category. Now, if we want a binary product of $ X, Y : \Alg T $ and we take the product of $ (T, X) $ and $ (T, Y) $ in the category of all algebras, we end up with $ (T \times T, X \times Y) $, or $ X \times Y : \Alg {T \times T} $, even though we would like to have $ (T, X) $ for some $ X : \Alg T $. It turns out that here we need the fact that algebras and presheaves are fibered over algebraic theories, and adding the following lemma to the library gave the final brick for showing show that all categories discussed here have limits:
\begin{lemma}
  For a displayed category $ D $ over a category $ C $, suppose that $ C $ has binary products, that $ D $ creates binary products, and that $ D $ is a fibration. Then the fiber categories $ D_X $ have binary products.
\end{lemma}
\begin{proof}
  Take $ \overline X_1, \overline X_2 : D_X $. We have a product $ \overline X_1 \times \overline X_2 : D_{X \times X} $ with projections $ \overline \pi_i : \overline X_1 \times \overline X_2 \to_{\pi_i} \overline X_i $. For the diagonal morphism on $ X $, the fibration gives an object $ \overline Y $ and a a cartesian lift:
  \[ \overline f : \overline Y \to_{\langle \id X, \id X \rangle} \overline X_1 \times \overline X_2, \]
  so we have projections
  \[ \overline p_i = \overline f \cdot \overline \pi_i : \overline Y \to_{\id X} \overline X_i. \]
  In the diagram below, the first row is the fiber $ D_X $ and the second row is the fiber $ D_{X \times X} $.
  \begin{center}
    \begin{tikzcd}
      \overline X_i & \overline Y \arrow[d, "\overline f" description] \arrow[l, "\overline p_i"] & \overline Z \arrow[ll, "\overline g_i"', bend right] \arrow[l, "\overline h"] \arrow[ld, "{\langle \overline g_1, \overline g_2 \rangle}"] \\
      & \overline X_1 \times \overline X_2 \arrow[lu, "\overline \pi_i"] &
    \end{tikzcd}
  \end{center}
  Now, given some $ \overline g_1: \overline Z \to_{g_1} \overline X_1 $ and $ \overline g_2: \overline Z \to_{g_2} \overline X_2 $, we have
  \[ \langle \overline g_1, \overline g_2 \rangle : \overline Z \to_{\langle g_1, g_2 \rangle} \overline X_1 \times \overline X_2. \]
  Because $ \overline f $ is cartesian, we have a unique
  \[ \overline h: \overline Z \to_{\id X} \overline Y, \]
  such that
  \[ \overline h \cdot \overline f = \langle \overline g_1, \overline g_2 \rangle. \]
  Therefore,
  \[ \overline h \cdot \overline p_i = \overline h \cdot \overline f \cdot \overline \pi_i = \langle \overline g_1, \overline g_2 \rangle \cdot \overline \pi_i = \overline g_i, \]
  which shows that $ \overline Y $ is the product of $ \overline X_1 $ and $ \overline X_2 $ in $ D_X $.
\end{proof}

@@

\section{Quotients}\label{sec:quotients}
This project used quotients twice, once explicitly and once implicitly. The first occurrence of quotients was the construction of the `extension of scalars' functor $ f_*: \RAct N \to \RAct M $ from Lemma \ref{lem:scalar-extension}. Here, the formalization worked directly with quotients over a relation. However, the relation given by $ R : (x n, m) \sim (x, f(n) \cdot m) $ for all $ a $, is reflexive and transitive, but not necessarily symmetric. Therefore, we first take the `equivalence closure' $ \overline R $ of $ R $: the smallest transitive and symmetric relation that contains $ R $. This is then used for the quotient. To construct functions from this, like the right action of the monoid $ M^\prime $, we use the universal property of the quotient: a function from $ X \times M^\prime $ that sends `related' elements $ (x_1, m_1) $ and $ (x_2, m_2) $ to the same element gives a function from the quotient $ X \times M^\prime / \sim $. However, the quotient is taken over $ \overline R $ instead of $ R $, so it is no longer possible to assume that $ x_2 = x_1 n $ and $ m_1 = f(n) \cdot m_2 $ for some $ n $. In the case of the right $ M^\prime $-action, it sufficed to add the following lemma to the library
\begin{lemma}
  For a relation $ S $ on a set $ A $ and a relation $ T $ on a set $ B $, if for $ f: A \to B $, we have that $ S(a_1, a_2) $ implies $ T(f(a_1), f(a_2)) $ for all $ a_1, a_2 : A $, then $ \overline S(a_1, a_2) $ implies $ \overline T(f(a_1), f(a_2)) $ for all $ a_1, a_2 $.
\end{lemma}
However, overall, working with these quotients in an elementary way was quite a hassle.

The second occurrence of quotients was in the proof of Corollary \ref{cor:karoubi-presheaf}, or the lemma that it is based on. Initially, the statement of the lemma was only about $ \SET $, stating that $ \iota_C \bullet - : [\overline C, \SET] \to [C, \SET] $ was an adjoint equivalence. The proof used coequalizers of functors, like in Lemma \ref{lem:karoubi-equivalence}. Note that coequalizers of set-valued functions are given pointwise by coequalizers in $ \SET $, which are formalized using quotients. Because of this, when simplifying, rocq usually reduced the goal to elementary statements about quotients, which somewhat obfuscated the relatively simple nature of the construction. Luckily, the construction could be made agnostic to the specific implementation of coequalizers, by introducing a variable
\[ \mathtt{H : Coequalizers HSET} \]
and using this instead of our known implementation. Since the proofs did not use any elements of the sets in question, it was then trivial to generalize to any category with coequalizers. It turned out that the construction was a special version of left Kan extension, which uses general colimits, and using the existing definition of Kan extension, I was able to considerably shorten the proof.

### 2024-12-09 800-conclusion.tex
Improve notation and wording

Also, since we saw that in univalent foundations, there are two nonequivalent definitions $ \bar \C $ and $ \hat \C $ for the category of retracts, it would be interesting to see how well Scott's and Taylor's proofs about $ \bar \C $, can be made to work on its Rezk completion $ \hat \C $. More generally, note that the Karoubi envelope is not specific to the material in this thesis. For example, one way to construct a cartesian closed category is by taking the Karoubi envelope of a `semi cartesian closed category', and the cartesian closed structure on $ \R $ in Scott's representation theorem can be viewed as a special case of this \autocite{hayashi-1985-semifunctors}. In another direction, the category of smooth manifolds can be constructed as the Karoubi envelope of the category $ \C $ of the open subsets of all Euclidean spaces, with smooth maps between them \autocite[][267]{Lawvere-Karoubi}. For such classical results about the Karoubi envelope, it would be interesting to study how they hold up in univalent foundations for the different choices $ \overline \C $ and $ \hat \C $ of `the Karoubi envelope'.

>>>>

Also, since we saw that in univalent foundations, there are two nonequivalent definitions $ \overline \C $ and $ \widehat \C $ for the category of retracts, it would be interesting to see how well Scott's and Taylor's proofs about $ \overline \C $, can be made to work on its Rezk completion $ \widehat \C $. More generally, note that the Karoubi envelope is not specific to the material in this thesis. For example, one way to construct a cartesian closed category is by taking the Karoubi envelope of a `semi cartesian closed category', and the cartesian closed structure on $ \R $ in Scott's representation theorem can be viewed as a special case of this \autocite{hayashi-1985-semifunctors}. In another direction, the category of smooth manifolds can be constructed as the Karoubi envelope of the category $ \C $ of the open subsets of all Euclidean spaces, with smooth maps between them \autocite[][267]{Lawvere-Karoubi}. For such classical results about the Karoubi envelope, it would be interesting to study how they hold up in univalent foundations for the different choices $ \overline \C $ and $ \widehat \C $ of `the Karoubi envelope'.
